---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, echo=TRUE}
library(tidyverse)
data <- read.csv("activity.csv")
```



## Mean total number of steps taken per day
```{r, echo=TRUE}
options(scipen=999)
ds <- data %>% group_by(date) %>% summarise(daily = sum(steps))
ms <- round(mean(ds$daily, na.rm=TRUE), 0)
md <- round(median(ds$daily, na.rm=TRUE), 0)
hist(ds$daily, main="Histogram of daily steps totals", xlab="Daily steps")
```

The mean of the daily steps is `r ms` and the median is `r md`.


## What is the average daily activity pattern?
```{r, echo=TRUE}
dint <- data %>% group_by(interval) %>% summarise(avg=mean(steps, na.rm=TRUE))
intmax <- dint[which.max(dint$avg),]$interval
plot(dint$interval, dint$avg, type="l", main="Time series plot of average steps per interval", xlab="Interval (5-minute)", ylab="Average steps")
```

The 5-minute interval which contains the maximum number of steps averaged across all the days is `r intmax`.

## Imputing missing values
Missing values will be the interval averaged ones calculated above.
```{r, echo=TRUE}

```


## Are there differences in activity patterns between weekdays and weekends?
