---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, echo=TRUE}
library(tidyverse)
library(timeDate)
library(lattice)
data <- read.csv("activity.csv")
```



## Mean total number of steps taken per day
```{r, echo=TRUE}
options(scipen=999)
ds <- data %>% group_by(date) %>% summarise(daily = sum(steps))
ms <- round(mean(ds$daily, na.rm=TRUE), 1)
md <- round(median(ds$daily, na.rm=TRUE), 1)
hist(ds$daily, main="Histogram of daily steps totals", xlab="Daily steps")
```

The mean of the daily steps is `r ms` and the median is `r md`.


## What is the average daily activity pattern?
```{r, echo=TRUE}
dint <- data %>% group_by(interval) %>% summarise(avg=mean(steps, na.rm=TRUE))
intmax <- dint[which.max(dint$avg),]$interval
plot(dint$interval, dint$avg, type="l", main="Time series plot of average steps per interval", xlab="Interval (5-minute)", ylab="Average steps")
```

The 5-minute interval which contains the maximum number of steps averaged across all the days is `r intmax`.

## Imputing missing values
Missing values will be the interval averaged ones calculated above.
```{r, echo=TRUE}
mv <- sum(is.na(data$steps))
cdata <- data.frame(data)
for (i in 1:length(cdata$steps)) {
  if (is.na(cdata$steps[i])) {
    cdata$steps[i] <- dint$avg[which(dint$interval == cdata$interval[i])]
  }
}
cds <- cdata %>% group_by(date) %>% summarise(daily = sum(steps))
cms <- round(mean(cds$daily), 1)
cmd <- round(median(cds$daily), 1)
hist(cds$daily, main="Histogram of NA corrected daily steps totals", xlab="Daily steps")
```

There are `r mv` missing values in the step data. I replaced the missing values with the average number of steps for the specific intervals calculated across all the days.

For this corrected data, the new mean of the daily steps is `r cms` and the new median is `r cmd`.  There is little impact on the mean and median locations by making these changes and the distribution is unchanged.


## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
cdata$when <- factor(isWeekend(as.POSIXct(cdata$date)), labels=c("Weekend", "Weekday"), levels=c(TRUE,FALSE))
by_group <- cdata %>% group_by(when, interval)
setint <- summarise(by_group, avg=mean(steps, na.rm=TRUE))
xyplot(avg ~ interval | when, data = setint, layout=c(1,2), type = "l", xlab = "Interval (5 min)", ylab = "Steps")
```


